Make any stupid plot you want

### 2021-05-05 draft/sections/intro.tex
data section revise

\subsection{Structure}
\label{sec:intro:structure}

\begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/intro/equivariant.png}
        \caption{The table (a) lists precipitation in millimeters in blue. These values are converted to  inches, in green, by dividing the millimeter values by 25.4. This conversion is a scaling action, represented by the orange arrows in (a). The \textit{equivariant} action is the scaling in the timeseries (b), which is represented by the orange fill between the green and blue lines. In (c), the scaling is illustrated in the different millimeter and inches y axis labeling of data that appears identical since the distances were scaled by a constant factor that is now incorporated in the labeling.}
      \label{fig:intro:equivariance}
\end{figure}
The components of a visual representation were first codified by Bertin\cite{bertinSemiologyGraphicsDiagrams2011a}, who introduced a notion of structure preservation that we formally describe in terms of \textit{equivariance} and \textit{continuity}. Bertin proposes that there are categories of visual encodings-such as position, shape, color, and texture-that preserve the properties of the measurement type, quantitative or qualitative, of the encoded data. For example, in \autoref{fig:intro:equivariance}, the blue precipitation data in millimeters in the table (a) is converted to inches. This scaling action is represented by the green arrows that translate the precipitation into the scaled values in green. For this visualization to be equivariant, this same scaling factor must be present in the timeseries representation of the data (b). The precipitation in green is an \textit{equivariant} scaling of the precipitation in blue, meaning that the y-values of the green line is $\frac{1}{25.4}$ that of the y-values of the blue precipitation timeseries. The orange fill in the timeseries (b) is the scaling equivalent to the arrows in the table (a). By definition, the distances in the millimeter data were scaled equally \cite{stothersSimilarityGroup} when converted to inches, which means the shapes of the graphs are equivalent when plotted against y-axes that preserve relative distance (c). This visualization is also equivariant to the table (a), but this is indicated through labeling rather than the line plots. The idea of equivariance is formally defined as the mapping of a binary operator from the data domain to the visual domain in Mackinlay's \textit{A Presentation Tool}(APT) model\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomaticDesignGraphical1987}. The algebraic model of visualization proposed by Kindlmann and Scheidegger uses equivariance to refer generally to invertible binary transformations\cite{kindlmannAlgebraicProcessVisualization2014}, which are mathematical groups \cite{shadrachIntroductionGroups2017}. Our model defines \textit{equivariance} in terms of monoid actions, which are a more restrictive set than all binary operations and more general than groups. As with the algebraic model, our model also defines structure preservation as commutative mappings from data space to representation space to graphic space, but our model uses topology to explicitly include continuity.

@@

Most information visualization software design patterns are tuned to specific data structures and domains, as categorized by Heer and Agrawala\cite{HeerSoftware2006}. For users who generally work in one domain--such relational tables (\autoref{fig:intro:table}), images (\autoref{fig:intro:image}), or graphs (\autoref{fig:intro:graph})--well defined data space (and corresponding visual space\cite{chiTaxonomyVisualizationTechniques2000}) often yields a coherent user experience\cite{heerDeclarative2010}. This coherant experience is often not extensible; developers who want to build new visualizations on top of these libraries must work around the existing assumptions, sometimes in ways that break the architecture model of the libraries. For example tools influenced by APT that assume that data is a relational table integrate computation into the visualization pipeline. This is a wide array of tools, including Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} and the Grammar of Graphics\cite{wilkinsonGrammarGraphics2005} inspired ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, protovis\cite{bostockProtoviz2009}, vega\cite{satyanarayanDeclarativeInteractionDesign2014} and altair\cite{vanderplasAltairInteractiveStatistical2018}. Since these libraries represent data as a table, and computations on tables are well defined\cite{ullmanFirstCourseDatabase2008}, these libraries implement computation as part of the visualization process. This is also true of tools that support images, such as the biology oriented ImageJ\cite{schneiderNIHImageImageJ2012} and  Napari\cite{nicholas_sofroniew_2021_4533308} or the digital humanities ImageJ macro ImagePlot\cite{studiesCulturevisImageplot2021}. While these tools often have some support for visualizing non image components of the data, the architecture is oriented towards building plugins into the existing system \cite{WritingPlugins} where the image is the core data structure. Tools like Gephi\cite{bastianGephiOpenSource2009}, Graphviz\cite{ellsonGraphvizOpenSource2002}, and Networkx\cite{HagbergExploringNetwork2008} are used to  visualize and manipulate graphs. As with tables and images, extending network libraries to work with other types of data either requires breaking the libraries internal model of how data is structured and what data transformations are allowed or implementing a model for other types of data structures alongside the network model. Our model aims to identify which computations are manipulations of the data and which are necessary for the visual encoding; for example data is often aggregated for a bar chart but it is not required, while a boxplot by definition requires computing distrobution statistics\cite{wickham40YearsBoxplots2011}. Disentangling the computation from the visual transforms allows us to determine whether the visualization library should implement computations or if they should be implemented in data space.

>>>>

\subsection{Structure}
\label{sec:intro:structure}
\begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/intro/equivariant.png}
        \caption{The table (a) lists precipitation in millimeters in blue. These values are converted to  inches, in green, by dividing the millimeter values by 25.4. This conversion is a scaling action, represented by the orange arrows in (a). The \textit{equivariant} action is the scaling in the timeseries (b), which is represented by the orange fill between the green and blue lines. In (c), the scaling is illustrated in the different millimeter and inches y axis labeling of data that appears identical since the distances were scaled by a constant factor that is now incorporated in the labeling.}
      \label{fig:intro:equivariance}
\end{figure}
The components of a visual representation were first codified by Bertin\cite{bertinSemiologyGraphicsDiagrams2011a}, who introduced a notion of structure preservation that we formally describe in terms of \textit{equivariance} and \textit{continuity}. Bertin proposes that there are categories of visual encodings-such as position, shape, color, and texture-that preserve the properties of the measurement type, quantitative or qualitative, of the encoded data. For example, in \autoref{fig:intro:equivariance}, the blue precipitation data in millimeters in the table (a) is converted to inches. This scaling action is represented by the green arrows that translate the precipitation into the scaled values in green. For this visualization to be equivariant, this same scaling factor must be present in the timeseries representation of the data (b). The precipitation in green is an \textit{equivariant} scaling of the precipitation in blue, meaning that the y-values of the green line is $\frac{1}{25.4}$ that of the y-values of the blue precipitation timeseries. The orange fill in the timeseries (b) is the scaling equivalent to the arrows in the table (a). By definition, the distances in the millimeter data were scaled equally \cite{stothersSimilarityGroup} when converted to inches, which means the shapes of the graphs are equivalent when plotted against y-axes that preserve relative distance (c). This visualization is also equivariant to the table (a), but this is indicated through labeling rather than the line plots. The idea of equivariance is formally defined as the mapping of a binary operator from the data domain to the visual domain in Mackinlay's \textit{A Presentation Tool}(APT) model\cite{mackinlayAutomatingDesignGraphical1986, mackinlayAutomaticDesignGraphical1987}. The algebraic model of visualization proposed by Kindlmann and Scheidegger uses equivariance to refer generally to invertible binary transformations\cite{kindlmannAlgebraicProcessVisualization2014}, which are mathematical groups \cite{shadrachIntroductionGroups2017}. Our model defines \textit{equivariance} in terms of monoid actions, which are a more restrictive set than all binary operations and more general than groups. As with the algebraic model, our model also defines structure preservation as commutative mappings from data space to representation space to graphic space, but our model uses topology to explicitly include continuity.

@@

Most information visualization software design patterns are tuned to specific data structures and domains, as categorized by Heer and Agrawala\cite{HeerSoftware2006}. For users who generally work in one domain--such relational tables (\autoref{fig:intro:table}), images (\autoref{fig:intro:image}), or graphs (\autoref{fig:intro:graph})--well defined data space (and corresponding visual space\cite{chiTaxonomyVisualizationTechniques2000}) often yields a coherent user experience\cite{heerDeclarative2010}. This coherant experience is often not extensible; developers who want to build new visualizations on top of these libraries must work around the existing assumptions, sometimes in ways that break the architecture model of the libraries. For example tools influenced by APT that assume that data is a relational table integrate computation into the visualization pipeline. This is a wide array of tools, including Tableau\cite{StoltePolaris2002,hanrahanVizQL2006,MackinlayShowme2007} and the Grammar of Graphics\cite{wilkinsonGrammarGraphics2005} inspired ggplot\cite{wickhamGgplot2ElegantGraphics2016a}, protovis\cite{bostockProtoviz2009}, vega\cite{satyanarayanDeclarativeInteractionDesign2014} and altair\cite{vanderplasAltairInteractiveStatistical2018}. Since these libraries represent data as a table, and computations on tables are well defined\cite{ullmanFirstCourseDatabase2008}, these libraries implement computation as part of the visualization process. This is also true of tools that support images, such as the biology oriented ImageJ\cite{schneiderNIHImageImageJ2012} and  Napari\cite{nicholas_sofroniew_2021_4533308} or the digital humanities ImageJ macro ImagePlot\cite{studiesCulturevisImageplot2021}. While these tools often have some support for visualizing non image components of the data, the architecture is oriented towards building plugins into the existing system \cite{WritingPlugins} where the image is the core data structure. Tools like Gephi\cite{bastianGephiOpenSource2009}, Graphviz\cite{ellsonGraphvizOpenSource2002}, and Networkx\cite{HagbergExploringNetwork2008} are used to  visualize and manipulate graphs. As with tables and images, extending network libraries to work with other types of data either requires breaking the libraries internal model of how data is structured and what data transformations are allowed or implementing a model for other types of data structures alongside the network model. Our model aims to identify which computations are manipulations of the data and which are necessary for the visual encoding; for example data is often aggregated for a bar chart but it is not required, while a boxplot by definition requires computing distribution statistics\cite{wickham40YearsBoxplots2011}. Disentangling the computation from the visual transforms allows us to determine whether the visualization library should implement computations or if they should be implemented in data space.

### 2021-05-05 draft/sections/math.tex
data section revise

\begin{figure}[H]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/monoid_total.png}
        \caption{}
        \label{fig:math:data:monoid:total}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=1\textwidth]{figures/math/monoid_partial.png}
        \caption{}
        \label{fig:math:data:monoid:partial}
    \end{subfigure}
    \caption{The action $\bullet$ in \autoref{fig:math:data:monoid:total} is the arrows from the partial order diagram of weather states on the left to the diagram of weather states on the right. The action maps the weather states to themselves, thereby preserving the ordering defined by the monoid $\ast$ on both sides of the action. We conjecture that the action in \autoref{fig:math:data:monoid:partial} is monotonic when the partial ordering of the weather states is the same as the partial order of the elements they are mapped to. Given $sun \geq clouds\geq rain,\, lightining$ on the right, we conjecture that the action  $sun,\, clouds \rightarrow clouds$, and $rain\, lightining \rightarrow rain$ is structure preserving such that the relative ordering of elements is the same as the elements they are mapped to.}
    \label{fig:math:data:monoid}
\end{figure}

@@

As with the fiber \dfiber\, the total monoid space \monoid\ is the cartesian product
\begin{equation}
    \label{eq:math:data:monoid:decompose}
\monoid= \monoid_{0} \times \ldots \times \monoid_{i}\times \ldots \times\ldots \monoid_{n}
\end{equation}
of each monoid $\monoid_{i}$ on $\dfiber_{i}$.  The monoid is also added to the specification of the fiber $(\fname_i,\, \ftype_i,\, \fttype\, \monoid_i)$

\subsubsection{Continuity of the Data $K$} 
\label{sec:math:data:base}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/k_different_types.png}
    \caption{The topological base space \dbase\ encodes the continuity of the data space, for example if the data is discrete points or lies on a plane or a sphere}
    \label{fig:math:data:base:types}
\end{figure}
The base space \dbase\ acts as an indexing space, as emphasized by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, to express how the records in \dtotal\ are connected to each other. As shown in \autoref{fig:math:data:base:types}, 
\dbase\ can have any number of dimensions and can be continuous or discrete. The base space also does not describe anything about the dataset besides the continuity. While the base space may have components to identify the continuity, such as \textit{time, latitutde, longitude}, these labels are indexed into from \dbase\ the same as any other component. This is similar to the notion of structural \textit{keys} with associated \textit{values} proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, but our model treats keys as a pure reference to topology. Decoupling the keys from their semantics allows the metadata to be altered; this provides for coordinate agnostic representation of the continuity and facilitates encoding of data where the independent variable may not be clear. For example the amount of snow on the ground is dependent on time of day and how much snow has fallen, and changing the coordinate system or time resolution should have no effect on how the records are connected to each other.

\begin{figure}[H]
    \includegraphics[width=\linewidth]{figures/math/k_qspace.png}
    \caption{The total space \dtotal\ is divided into fiber segments \dfiber. The base space \dbase\ acts as an index into the records in the fibers, such that every point \dbasepoint\ has a corresponding fiber $\dfiber_{\dbasepoint}$.}
    \label{fig:math:data:base:div}
\end{figure}

Formally \dbase\ is the quotient space \cite{QuotientSpaceTopology2020} of \dtotal\, meaning it is the finest space\cite{aurouxMath131Introduction} such that every $\dbasepoint \in \dbase$ has a corresponding fiber $\dfiber_k$\cite{QuotientSpaceTopology2020}. In \autoref{fig:math:data:base:div}, \dtotal\ is a rectangle divided by vertical fibers \dfiber, so the minimal \dbase\ for which there is always a mapping $\pi: \dtotal\rightarrow \dbase$ is the closed interval $\left[0,1\right]$.

As with \autoref{eq:math:data:fiber:decompose} and \autoref{eq:math:data:monoid:decompose}, we can decompose the total space into component bundles $\pi:\dtotal_i\rightarrow \dbase$ where
\begin{equation}
    \label{eq:math:data:base:decompose}
    \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
\end{equation}
such that \(\monoid_{i}\) acts on component bundle \(\dtotal_i\). The \dbase\ remains the same because the connectivity of records does not change just because there are fewer components in each record. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit.
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/base.png}
    \caption{The fiber bundles in (a) and (b) encode the two component dataset from \autoref{fig:math:data:fiber}, with \textit{(time, precipitation)} components, as having different continuities. The fiber bundle with discrete continuity (a) encodes the dataset as being a set of discrete records. The fiber bundle over the continuous interval \dbase\ (b) encodes the records as if they were sampled from a 1D continuous space.} 
    \label{fig:math:data:base}
\end{figure}

The datasets in \autoref{fig:math:data:base} have the same fiber of (precipitation, time). The dot represents a discrete base space \dbase, meaning that every dataset encoded in the fiber bundle has discrete continuity. The line is a representation of a 1D continuity, meaning that every dataset in the fiber bundle is 1D continous. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit. The explicit topology is a concise way of distinguishing visualizations that appear identical, for example heatmaps and images.

\subsubsection{Data \dsection}
\label{sec:math:data:section}
While the projection function $\pi:\dtotal \rightarrow\dbase$ ties together the base space \dbase\ with the fiber \dfiber, a section $\dsection: \dbase\rightarrow \dtotal$ encodes a dataset. A section function takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For example, in the special case of a table \cite{spivakSIMPLICIALDATABASES}, \dbase\ is a set of row ids, \dfiber\ is the columns, and the section \dsection\ returns the record \delement\ at a given key in \dbase. For any fiber bundle, there exists a map
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
 such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section is 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. Because we can decompose the bundle and the fiber (\autoref{eq:math:data:base:decompose}, \autoref{eq:math:data:fiber:decompose}), we can decompose \dsection\ as 
\begin{equation}
\label{eq:math:data:section:decompose}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{equation}
where each section $\dsection_i$ maps into a record on a component $\dfiber_i \in \dfiber$. This allows for accessing the data component wise in addition to accessing the data in terms of its location over \dbase.

\begin{figure}[H]
    \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
    \caption{ Fiber (time, precipitation) with an interval \dbase\ basespace. The sections $\dsection^{(1)}$ and $\dsection^{(2)}$ are constrained such that the time variable must be monotonic, which means each section is a timeseries of precipitation values. They are included in the global set of sections  $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$}
    \label{fig:math:data:sections}
\end{figure}

@@

\subsubsection{Sheafs}
\label{sec:math:data:sheaf}
Many types of dynamic visualizations require evaluating sections on different subspaces of \dbase, an a sheaf, denoted $\mathcal{O}$ provides a way to do so. A sheaf is a mathematical structure for defining collections of objects\cite{ghristElementaryAppliedTopology2014, ghristHomologicalAlgebraData2018,urbanikBriefIntroductionSchemes} on mathematical spaces. On the fiber bundle \dtotal, we can describe a sheaf as the collection of local sections $\iota^{*}\dsection$
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}[column sep=huge]
        \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
        U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
which are sections of \dtotal\ pulled back over local neighborhood $U \subset \dtotal$ via the inclusion map \(\iota:\dtotal\rightarrow U\). The collation of sections enabled by sheafs is necessary for navigation techniques such as pan and zoom\cite{NekrasovskiEvaluationPanZoom2006} and dynamically updated visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

\subsubsection{Applications to Data Containers}
This model provides a common formalism for widely used data containers without sacrificing the semantic structure embedded in each container. For example, the section can be any instance of a univariate numpy array\cite{harris2020array} that stores an image.  This could be a section of a fiber bundle where \dbase\ is a 2D continuous plane and the \dfiber\ is $(\reals^{3}, \reals, \reals)$ where $\reals^3$ is color, and the other two components are the x and y positions of the sampled data in the image. This position information is already implicitely encoded in the array as the index and the resolution of the image being stored.Instead of an image, the numpy array could also store a 2D discrete table. The fiber would not change, but the \dbase\ would now be 0D discrete points. These different choices in topology indicate, for example, what sorts of interpolation would be appropriate when visualizing the data.

There are also many types of labeled containers that can richly be described in this framework because of the schema like structure of the fiber. For example, a pandas series which stores a labeled list, or a dataframe\cite{jeff_reback_2020_3715232} which stores a relational table. A series could store the values of $\dsection^{(1)}$ and a second series could be  $\dsection^{(2)}$. We could also fatten the fiber to hold two precipitationure series, such that a section would be an instance of a dataframe with a time column and two precipitation columns. While the series and dataframe explicitly have a time index column, they are components in our model and the index is assumed to be data independent references such as hashvalues, virtual memory locations, or random number keys.

Where this model particularly shines are N dimensional labeled data structures. For example, an xarray\cite{hoyer2017xarray} data that stores precipitation field could have a \dbase\ that is a continuous volume and the components would be the precipitation and the time, latitude, and longitude the measurements were sampled at. A section can also be an instance of a distributed data container, such as a dask array \cite{rocklinDaskParallelComputation2015}. As with the other containers, \dbase\ and \dfiber\ are defined in terms of the index and dtypes of the components of the array. Because our framework is defined in terms of the fiber, continuity, and sections, rather than the exact values of the data, our model does not need to know what the exact values are until the renderer needs to fill in the image.

>>>>

\begin{figure}[H]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/monoid_total.png}
        \caption{}
        \label{fig:math:data:monoid:total}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=1\textwidth]{figures/math/monoid_partial.png}
        \caption{}
        \label{fig:math:data:monoid:partial}
    \end{subfigure}
    \caption{The action $\bullet$ in \autoref{fig:math:data:monoid:total} is the arrows from the partial order diagram of weather states on the left to the diagram of weather states on the right. The action maps the weather states to themselves, thereby preserving the ordering defined by the monoid $\ast$ on both sides of the action. We conjecture that the action in \autoref{fig:math:data:monoid:partial} is monotonic when the partial ordering of the weather states is the same as the partial order of the elements they are mapped to. Given $sun \geq clouds\geq rain,\, lightining$ on the right, we conjecture that the action  $sun,\, clouds \rightarrow clouds$, and $rain,\, lightining \rightarrow rain$ is structure preserving such that the relative ordering of elements is the same as the elements they are mapped to.}
    \label{fig:math:data:monoid}
\end{figure}

@@

As with the fiber \dfiber\, the total monoid space \monoid\ is the cartesian product
\begin{equation}
    \label{eq:math:data:monoid:decompose}
\monoid= \monoid_{0} \times \ldots \times \monoid_{i}\times \ldots \times\ldots \monoid_{n}
\end{equation}
of each monoid $\monoid_{i}$ on $\dfiber_{i}$.  The monoid is added to the specification of the fiber $(\fname_i,\, \ftype_i,\, \fttype\, \monoid_i)$

\subsubsection{Continuity of the Data $K$} 
\label{sec:math:data:base}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/k_different_types.png}
    \caption{The topological base space \dbase\ encodes the continuity of the data space. The table of discrete weather station records has discrete continuity such that each record maps to a single point (a). A gaussian has a value at all points along the interval x is sampled from and therefore has a 1D continuity (b). The globe has a value at all points (latitude, longitude) on the globe and therefore has 2D continuity (c). }
    \label{fig:math:data:base:types}
\end{figure}

The base space \dbase\ provides a way to explicitly encode the continuity of the data, as described in \autoref{sec:intro:data:continuity}. This explicit topology is a concise way of distinguishing between visualizations that appear identical but assume different continuity, for example heatmaps and images. The base space \dbase\ acts as an indexing space, as emphasized by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, to express how the records in \dtotal\ are connected to each other. As shown in \autoref{fig:math:data:base:types}, \dbase\ can have any number of dimensions and can be continuous or discrete. Formally \dbase\ is the quotient space \cite{QuotientSpaceTopology2020} of \dtotal\, meaning it is the finest space\cite{aurouxMath131Introduction} such that every $\dbasepoint \in \dbase$ has a corresponding fiber $\dfiber_k$\cite{QuotientSpaceTopology2020}.

\begin{figure}[H]
    \includegraphics[width=\linewidth]{figures/math/k_qspace.png}
    \caption{The total space \dtotal\ is divided into fiber segments \dfiber. The base space \dbase\ acts as an index into the records in the fibers, such that every point \dbasepoint\ has a corresponding fiber $\dfiber_{\dbasepoint}$. The projection map $\pi$ maps every fiber $\dfiber_{\dbasepoint}$ to a point $\dbasepoint \in \dbase$ in the base space.}
    \label{fig:math:data:base:div}
\end{figure}

In \autoref{fig:math:data:base:div}, \dtotal\ is a rectangle divided by vertical fibers \dfiber, so the minimal \dbase\ for which there is always a mapping $\pi: \dtotal\rightarrow \dbase$ is the closed interval $\left[0,1\right]$. While the total space \dtotal\ may have components in \dfiber\ that describe any given point $\dbasepoint \in \dbase$, such as \textit{time, latitutde, longitude}, these labels are indexed into from \dbase\ the same as any other components. In contrast to the structural \textit{keys} with associated \textit{values} proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, our model treats keys \dbasepoint\ as a pure reference to topology. Decoupling the keys from their semantics allows the components identifying the keys to be altered, which provides for coordinate agnostic representation of the continuity and facilitates encoding of data where the independent variable may not be clear. For example total rainfall is dependent on time of day and how much rain has already fallen, and changing the coordinate system or time resolution should have no effect on how the records are connected to each other.

As with \autoref{eq:math:data:fiber:decompose} and \autoref{eq:math:data:monoid:decompose}, we can decompose the total space into component bundles $\pi:\dtotal_i\rightarrow \dbase$ where
\begin{equation}
    \label{eq:math:data:base:decompose}
    \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
\end{equation}
such that the monoid \(\monoid_{i}\) acts on component bundle \(\dtotal_i\). The \dbase\ remains the same because the continuity of the data does not change just because there are fewer components in each record.

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/base.png}
    \caption{The fiber bundles in (a) and (b) encode the two component dataset from \autoref{fig:math:data:fiber}, with \textit{(time, precipitation)} components, as having different continuities. The fiber bundle with discrete continuity (a) encodes the dataset as being a set of discrete records. The fiber bundle over the continuous interval \dbase\ (b) encodes the records as if they were sampled from a 1D continuous space.} 
    \label{fig:math:data:base}
\end{figure}

The datasets in \autoref{fig:math:data:base} have the same fiber of (precipitation, time). The points (a) represent a discrete base space \dbase, meaning that every dataset encoded in the fiber bundle has discrete continuity. The line (b) is a representation of a 1D continuity, meaning that every dataset in the fiber bundle is 1D continuous. Explicitly encoding data continuity, for example that (a) has discrete continuity and (b) is 1D continuous, provides a means to explicitly specify the continuities visualization components must preserve.

\subsubsection{Data \dsection}
\label{sec:math:data:section}
While the projection function $\pi:\dtotal \rightarrow\dbase$ ties together the base space \dbase\ with the fiber \dfiber, a section $\dsection: \dbase\rightarrow \dtotal$ encodes a dataset. A section function takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For example, in the special case of a table \cite{spivakSIMPLICIALDATABASES}, \dbase\ is a set of row ids, \dfiber\ is the columns, and the section \dsection\ returns the record \delement\ at a given key in \dbase. For any fiber bundle, there exists a map
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
 such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section can be decomposed as 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. Because we can decompose the bundle and the fiber (\autoref{eq:math:data:base:decompose}, \autoref{eq:math:data:fiber:decompose}), we can decompose \dsection\ as 
\begin{equation}
\label{eq:math:data:section:decompose}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{equation}
where each section $\dsection_i$ maps into a record on a component $\dfiber_i \in \dfiber$. This allows for accessing the data component wise in addition to accessing the data in terms of its location over \dbase.

\begin{figure}[H]
    \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
    \caption{ Fiber (time, precipitation) with a 1D continuous \dbase\  defined on an interval $\left[0,n\right]$. The sections $\dsection^{(1)}$ and $\dsection^{(2)}$ are constrained such that the time variable must be monotonic, which means each section is a timeseries of precipitation values. They are included in the global set of sections  $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$}
    \label{fig:math:data:sections}
\end{figure}

@@

\subsubsection{Sheafs}
\label{sec:math:data:sheaf}
Dynamic visualizations require evaluating sections on different subspaces of \dbase; this can be achieved using a mathematical structure, called a sheaf $\mathcal{O}$, for defining collections of objects\cite{ghristElementaryAppliedTopology2014, ghristHomologicalAlgebraData2018,urbanikBriefIntroductionSchemes} on mathematical spaces.  On the fiber bundle \dtotal, we can describe a sheaf as the collection of local sections $\iota^{*}\dsection$
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}[column sep=huge]
        \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
        U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
which are sections of \dtotal\ pulled back over local neighborhood $U \subset \dtotal$ via the inclusion map \(\iota:\dtotal\rightarrow U\). The collation of sections enabled by sheafs is necessary for navigation techniques such as pan and zoom\cite{NekrasovskiEvaluationPanZoom2006} and dynamically updated visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

\subsubsection{Applications}
Using fiber bundles as our data abstraction allows the model to be applicable to a wide variety of fields and use cases. For example, the section can be any instance of a numpy array\cite{harris2020array} that stores an image, such as an image where the \dbase\ is a 2D continuous plane and the \dfiber\ is $(\reals^{3}, \reals, \reals)$. In this fiber, the $\reals^3$ components encode color, and the other two components are the x and y positions of the sampled data in the image. The continuity of the image is implicitly encoded in the array as the index, so the position components encode the resolution. Instead of an image, the numpy array could also store a 2D discrete table. The fiber may not change, but the \dbase\ would now be 0D discrete points. These different choices in topology indicate, for example, what sorts of interpolation would be appropriate when visualizing the data. Labeled containers can also be described in this framework because of the schema like structure of the fiber. One such example is a pandas series which stores a labeled list, another is a  dataframe\cite{jeff_reback_2020_3715232} which has the structure of a relational table. A series could store the values of $\dsection^{(1)}$ and a second series could be  $\dsection^{(2)}$, while a dataframe would have multiple components and each data frame would be a unique section \dsection. The ability to encode complexity in continuity and components is particulary beneficial when working with  N dimensional labeled data containers. For example, an xarray\cite{hoyer2017xarray} data cube that stores precipitation would be a section of a fiber bundle with a \dbase\ that is a continuous volume and components \textit{(time, latitude, longitude, precipitation)}. This section does not need to resolve to values immediately and instead can be an instance of a distributed data container, such as a dask array \cite{rocklinDaskParallelComputation2015}. Our model provides a common formalism for describing widely used data containers without sacrificing the semantic structure embedded in each container.
